Formation of $\cenn$ Molecules in Oxygen-rich Interiors of Type II Supernovae

### 2013-03-12 doc/submit/ms.tex
Update for the referees review.

The problem historically of condensing carbon SUNOCONs in the interiors of
expanding Type II supernovae is that most of the synthesized carbon is bathed
in more abundant oxygen and was long presumed to be oxidized by it. This
changed when \citet{1999Sci...283.1290C} introduced a kinetic theory by
which carbon can condense thermally in gas having more abundant oxygen.
That theory is noteworthy in having every chemical reaction from the initial
O $>$ C gas to macroscopic graphite included in a kinetic network, even though
most of the cross sections for those reactions and the rates of isomeric
transitions from linear carbon chains to carbon rings are not accurately known.
Their paper provided an {\em existence proof}
that carbon can condense in O-rich gas.
Its calculations showed that rare carbon-ring molecules were not prevented from
formation despite rapid oxidation of their molecular precursors, although
oxidation does greatly reduce the abundance of those $\cenn$ chains.
That initial
theory has been amplified during the past decade by several studies
\citep{2001ApJ...562..480C,2003ApJ...594..312D,2006ApJ...638..234D,
2009ApJ...703..642C,2010ApJ...713....1C,2011NewAR..55..155C,
2013ApJ...762....5C}.

In basic agreement with this chemical model, \citet{2009ApJ...703..642C} raised
questions about the most effective way of producing $\ctwo$\ molecules,
which is the
first step to linear carbon chains. Using a large network of chemical
reactions, \citet{2009ApJ...703..642C} discovered that the neutral-neutral
reaction \coctoctwo\
has a reasonably large cross section at temperatures near
5000K despite being endothermic by 4.8eV. Although the CO target abundance for
production of $\ctwo$\ in this manner must be calculated, they stated that at
$T=5000$ K the rate factor
$\langle \sigma v \rangle = k_{CO,C} = 8.6 \times 10^{-14}$ cm$^3$ s$^{-1}$
created $\ctwo$\ much more
rapidly than the slow radiative reaction \twoctoctwo.
Sensing its significance,
they claimed in their list of conclusions: ``2. A new pathway to the formation
of carbon chains is active in the O-rich mass zone of the unmixed ejecta and
is identified as the CO conversion to $\ctwo$\ via collisions with C.''  They were
unable to evaluate its full quantitative consequences because their reaction
network terminated at $\cthree$\
and did not include the formation of linear carbon
chains $\cenn$
and their isomerization to carbon rings. Doing so is one goal of our
study. But the overriding goal is to point out large abundance differences
between the computations of these two groups and to resolve them if possible.

Several important consequences for astronomy depend on careful
analysis of this claim that the production of $\ctwo$\ owing to the reaction
\twoctoctwo, as \citet{1999Sci...283.1290C,2001ApJ...562..480C}
had done, severely
underestimates the abundances of linear carbon chains because much more $\ctwo$\
is made earlier \citep{2010ApJ...713....1C} near $T = 5000$ K. Testing this
competition is essential for understanding of dust condensation within
supernovae. Dust created by supernova expansions is currently under intense
study owing to three types of astronomical observations: (1) of dust observed
in single supernova remnants; (2) of dust observed in early low-metallicity
galaxies; (3) of supernova-condensed carbon dust (SUNOCONs) extracted from
meteorites. Each of these topics depends sensitively on how much carbon,
both numbers and sizes, is able to condense in cooling SNII interiors.
Therefore we study this competition carefully.

Crucial to this task is the lifetime $\tau_{CO}$ of CO molecules. Their thermal
dissociation lifetime $\tau_\gamma$
is calculated from detailed balance with the
radiative association reaction, as in Section 2.1 of
\citet{2001ApJ...562..480C}.
Because of the large 11.1eV binding energy of the CO molecule,
$\tau_\gamma$ is very
temperature sensitive.  For readers' numerical ease we will tabulate
$\tau_\gamma$ at
selected key temperatures in Table \ref{tab:quantities} to follow.
The flux and spectrum of newly
injected Compton electrons \citep{1991ApJ...375..221C} yields the lifetime
$\tau_e$ of CO
molecules against inelastic scattering dissociation by Compton electrons
caused by the $^{56}$Co
radioactivity. In a gas of pure CO the mean energy per ion
pair is defined as the energy of primary electrons divided by the number of
pairs produced. \citet{1994ApJ...435..909L}
calculated the mean energy per ion pair
in pure CO gas and obtained the result $\Delta E = 32.3$ eV
deposited per dissociated CO
pair, which agrees well with the measurement of 32.2 eV by
\citet{1968...Klots}.
So the efficiency of energetic electrons for dissociating CO seems well
established. A lifetime $\tau_{CO}$ near one week is typical in SN 1987A, but
would be longer in SNII synthesizing less $^{56}$Co and at times greater than
about 8 months.

@@

Because our goal is to study the chemistry, we can take a very simple physical
model of the expansion; namely, temperature
$T = 3800 K / (t / 100 {\rm d}) =
3.30 \times 10^{10} {\rm K} / t({\rm s})$,
where t is the time elapsed since core collapse and $n_O = 10^{10}$ cm$^{-3}$
and $n_C = 10^9$ cm$^{-3}$
at the starting time $t = t(6000 {\rm K})$ for our chemical network at
$T=6000$ K. From our model choice for $T(t)$ we get
$t(6000K) = 5.47 \times 10^6$s. At
subsequent times $n_O(t) = 10^{10}{\rm cm}^{-3}(t/t(6000 {\rm K}))^{-3}$
owing to homologous expansion.
Let $N$ be the number in a comoving, expanding, initially 1 cm$^3$
volume at 6000 K
that held initially $n_O=10^{10}$ cm$^{-3}$ and $n_C = 10^9$ cm$^{-3}$.
The only change of $N$ during
expansion occurs through chemical reactions. $N$ may be expressed as atom
fraction $Y$ of the initial total number
$N_O + N_C = 1.10 \times 10^{10}$ atoms. We
intentionally choose an O-rich interior having $N_O/N_C = 10$,
so that no carbon at
all would be able to condense if that interior were governed by chemical
equilibrium.

We use an intentionally limited set of chemical species because our goal is
to study carbon chemistry: C, O, CO, $\ctwo$, $\cthree$, ... $\ceight$\ and
$\ceightr$, the ringed isomer of $\ceight$, to which we give a lifetime
$\tau_8^r = 10$ s against thermal isomeric transition
from linear $\ceight$ to ringed $\ceightr$.
Those ringed molecules are the seeds for carbon
growth because their oxidation rates are much smaller than oxidation rates of
linear $\cenn$
whereas their C capture rates are fast. We take ringed $\ceightr$ to be
indestructible, simply integrating their rate of production.
We cannot allege that $\ceight$\ is the key isomerizing chain, nor that
$\tau_8^r = 10$ s is its lifetime, but those assumptions are adequate for
comparing models for formation for $\ctwo$.
Our strategy is
to compute the number $\ceightr$ remaining after expansion from two differing
sources of $\ctwo$\ initiating the linear $\cenn$
chains. Those two source reactions are
\begin{itemize}
\item \twoctoctwo\ \citep{1999Sci...283.1290C}.
\item \coctoctwo\ \citep{2009ApJ...703..642C,2010ApJ...713....1C}.
\end{itemize}
We take our reaction rates from the rate tables in
\citet{2009ApJ...703..642C,2010ApJ...713....1C}.
The rate for the second reaction is indeed
much greater than that of the first reaction above $T=3000$ K;
but it becomes increasingly the smaller of the two below $T=2500$ K.
We include the thermal inverse reaction of every reaction,
which we calculate as in Eq.(3) of \citet{2001ApJ...562..480C}.
See Table 1 below for sample thermal dissociation rates for CO molecules.
We also include in our network dissociation of CO by Compton electrons.
CO is the only molecule for which electron dissociation can be the dominant
destruction mechanism. For that rate we use
$\tau_e (s) = 10^5 \exp( (t-10^6{\rm s}) / 111 {\rm d} )$,
not fit to any specific model but with plausibility for SN1987A.
It lengthens to $\tau_e = 10^6$ s near eight months.
Because of their helpful cataloging of reaction rate tables,
and in order to compare results without rate differences, we take the
rates as given in the tables of
\citet{2009ApJ...703..642C,2010ApJ...713....1C}.
As one example, we take the rate for \twoctoctwo\
as the rate given by RA4 in Table 5 in \citet{2009ApJ...703..642C}, p. 649.

@@

Evidently the abundance of CO within that zone during expansion attempts to
balance these creation and destruction effects, doing so exactly at the
time of maximum $n_{CO}$. \citet{2013ApJ...762....5C}
has discussed the shape of the function
$n_{CO}$(t). His Eq.(1), which is valid at constant density,
approximates the growth of $n_{CO}$ by its leading terms:
\begin{equation}
\frac{dn_{CO}}{dt} = n_C n_O k_{CO} - \frac{n_{CO}}{\tau_{CO}} = 0
\label{eq:dncodt}
\end{equation}
at $t = t_{max}$.
The maximum abundance reached by CO is in cgs units
\begin{equation}
n_{CO}(t_{max}) = n_C n_O k_{CO} \tau_{CO} 
\label{eq:comax}
\end{equation}

@@

Figure \ref{fig:yi}
shows the abundances $Y_i$ of each species in our small network as a
function of time $t-t(6000 {\rm K})$ after $T=6000$ K.
From Table \ref{tab:quantities},
start time is $t(6000K) =5.47 \times 10^6$ s.
To display each $Y_i$ on a figure with
reasonable ordinate resolution, we have scaled each $\cenn$\ by the factor
stated in the box. Many features are noteworthy: (1) C and O are constant
because their small depletion is negligible on the scale shown;
(2) maxima in CO and $\ctwo$\ occur at almost exactly the same time
$t-t(6000K)= 4 \times 10^6$ s, as was also seen in Fig. \ref{fig:ncoc2};
(3) $\cthree$\ has very small abundance, about $3 \times 10^{-8}$ of $\ctwo$
at the $\ctwo$\ maximum, although much later $\cthree$\ slowly grows modestly
relative to much more abundant $\ctwo$; (4) the rise shapes of $\cfour$
through $\ceight$\ are very similar because they are linked by a near steady
state; (5) the ringed carbon $\ceightr$\ has similar abundance shape vs. time,
but notice that it is actually much more abundant than $\cseven$ and
$\ceight$ because it accumulates from isomeric transitions of $\ceight$\
and unlike $\cenn$\ does not
suffer from fast oxidation \citep{1999Sci...283.1290C}. Each of these
features is understandable in terms of the flows into and out of each species.

One sees from Fig. \ref{fig:Yi} that the \coctoctwo\ reaction plays no
role in $\ceightr$\ production from the fact that $\ceightr$\ rises only for
$t - t(6000 {\rm K}) > 1.5 \times 10^7$ s, corresponding to $T < 2000$ K,
despite $\ctwo$\ and $\cthree$\ rising at much earlier times.
Whatsoever $\ctwo$\ is made earlier plays no role in $\ceightr$\ production
because there ejection of C atoms from $\cenn$\ at higher $T$ prevents the
flow to $\ceight$.  Only at $T < 2000$ K do those photodissociation reactions
become so slow that $\ceight$\ can grow, which it does from the $\ctwo$\
recently
formed by the \twoctoctwo\ reaction near 2000 K.  Fig. \ref{fig:Yi} also shows
that $Y(\ceightr)$
grows quickly to $10^{-16}$.  Since available C is $Y_C = 0.1$,
the ratio gives $10^{-15}$ C/ring.  This is enough to grow very large
graphite.

Figure \ref{fig:cn} shows the shape of $\cenn$\ vs. $n$ for $\cthree$\ to
$\ceight$\ at two different temperatures near 2000 K. As temperature declines,
the curve flattens because photoejection from $\cenn$\ by thermal photons
(or phonons) weakens. These abundance ratios are almost in a steady state,
but that steady state changes slightly as $T$ falls. These patterns show
almost equal values for $Y(\cthree)$ because that curve is relatively flat
with time near $t-t(6000K) = 10^7$ s (Fig. \ref{fig:yi}).
Table \ref{tab:flows} lists our computed reaction flows both in and out of
$\csix$\ at $T = 2031$ K and $T = 1802$ K.
The reactions are specified there in the
compact notation target(in,out)residual. Fig. \ref{fig:cn} and
Table \ref{tab:flows} are made to be studied together.
Note that the abundances of $\ceight$\ are very small:
$Y(\ceight)= 10^{-36}$ at $T = 1802$ K.
The isomeric transition from linear to
ringed $\ceight$, for which we estimate a lifetime $\tau = 10$ s,
provides the nucleations for graphite growth.

We call attention to several conclusions to be drawn from Table \ref{tab:flows}.
Firstly, the strongest flows by a wide margin involving $\csix$\ are
\cfivectocsix,
which is very fast ($k = 3 \times 10^{-10} {\rm cm}^3 {\rm s}^{-1}$)
because vibrational excitation of $\csix$
obviates the need for creating a photon \citep{1999Sci...283.1290C},
and its inverse reaction which ejects a C atom. Furthermore, those flows
are equal to three significant figures, illustrating the near steady
state of the abundance pattern. Secondly, thermal dissociation of $\csix$\ is
very much faster than its oxidation, showing the small effect of oxidation
on the abundance pattern. The destroying flows from $\csix$\ to $\cfive$
stand in the approximate ratio
${\rm C}_6(\gamma,{\rm C})/{\rm C_6}({\rm O},{\rm CO}) = 5 \times 10^5$
at $T = 2031$ K and $2 \times 10^4$ at $T = 1802$ K.
Dissociation by thermally excited vibrations dominates oxidation. What we here
label ($\gamma,{\rm C}$) is actually radiationless. Note that oxidation of
$\cenn$ molecules is not faster than C-ejection reactions. This unusual
situation occurs because the C-capture reactions proceed by exciting 
vibrations of the $\csix$ molecule. That residual vibrational energy makes
capture reactions fast rather than slow, but also makes the inverse C-ejection
reactions caused by thermally excited vibrations to be much faster. Detailed
balance gets the ratio right. The smaller value for that ratio at $T=1802$ K
occurs because oxidation maintains its effectiveness as $T$ drops but
thermal disruption does not, being much more temperature sensitive.
For this reason the abundance pattern is flatter and $\csix$\ is approximately
3000 times more abundant at $T=1802$ K than at $T=2031$ K.
It is for that reason that destruction flows in Table \ref{tab:flows}
for $\csix$\ are
larger at the smaller temperature.
The steady state shifts to new ratios as $T$ falls.
Thirdly, calculation of the rates will be illustrated for sake of clarity
by the flow
\csixotococfive.
From Fig. \ref{fig:cn}
one sees that at $T = 2031$ K the abundance $Y(\csix) = 10^{-33.3}$.
The value of $Y(O) = 10^{10} cm^{-3} / 1.1 \times 10^{10} cm^{-3} = 0.909$,
so the flow per atom per
second (e.g., Eq. \ref{eq:dycodt}) is
$dY/dt= Y(\csix)Y(O)n(2031K)k(\csix + {\rm O})$.
Expansion from $T=6000$ K to $T=2031$ K has diluted the total number
density $n$ to $1.1 \times 10^{10}{\rm cm}^{-3}
(2031 {\rm K}/6000 {\rm K})^3 = 4.27 \times 10^8{\rm cm}^{-3}$.
The fast oxidation reaction rate is
$k(\csix + {\rm O}) = 3 \times 10^{-10} {\rm cm}^3{\rm s}^{-1}$.
Gathering factors yields approximately $4 \times 10^{-35}$ per atom
per second
in good approximation to the Table \ref{tab:flows}
flow entry $3.61 \times 10^{-35}$ per atom per second.

@@

\begin{table}
\begin{center}
\caption{Values of $\tau_\gamma$(CO) and related quantities at
selected time $t$ and temperature $T$}
\label{tab:quantities}
\begin{tabular}{ccccccc}
\tableline\tableline
$t$ (s) & $t - t(6000 {\rm K})$ (s) & $\left(t / t(6000 {\rm K} \right)^{-3}$ & $T$ (K) & $\tau_\gamma$ (s) & $k_{CO}$ (cm$^{-3}$ s$^{-1}$) & $\ncogeq$ (cm$^{-3}$) \\
\tableline
$5.47 \times 10^6$ &
  0 &
  1 &
  6000 &
  $4.22 \times 10^{-2}$ &
  $3.13 \times 10^{-17}$ &
  $1.32 \times 10^1$ \\
$5.97 \times 10^6$ &
  $4.97 \times 10^5$ &
  0.769 &
  5500 &
  $3.54 \times 10^{-1}$ &
  $2.99 \times 10^{-17}$ &
  $6.27 \times 10^1$ \\
$6.57 \times 10^6$ &
  $1.09 \times 10^6$ &
  0.578 &
  5000 &
  $4.46 \times 10^{0}$ &
  $2.83 \times 10^{-17}$ &
  $4.24 \times 10^2$ \\
$7.30 \times 10^6$ &
  $1.82 \times 10^6$ &
  0.422 &
  4500 &
  $9.70 \times 10^{1}$ &
  $2.67 \times 10^{-17}$ &
  $4.61 \times 10^3$ \\
$8.21 \times 10^6$ &
  $2.74 \times 10^6$ &
  0.296 &
  4000 &
  $4.43 \times 10^{3}$ &
  $2.48 \times 10^{-17}$ &
  $9.67 \times 10^4$ \\
$9.38 \times 10^6$ &
  $3.91 \times 10^6$ &
  0.198 &
  3500 &
  $5.86 \times 10^{5}$ &
  $2.28 \times 10^{-17}$ &
  $5.26 \times 10^6$ \\
$1.09 \times 10^7$ &
  $5.47 \times 10^6$ &
  0.125 &
  3000 &
  $3.77 \times 10^{8}$ &
  $2.05 \times 10^{-17}$ &
  $1.21 \times 10^9$ \\
\tableline
\end{tabular}
\end{center}
\end{table}

@@

>>>>

Whether carbon dust can or can not condense in the interiors of expanding 
Type II supernovae is an important question for astronomy. Many have 
assumed, guided by chemical equilibrium, that carbon condensation can 
not occur unless carbon is more abundant than oxygen. That assumption 
limits C condensation to the He-burning shell of SNII. This rule-of-thumb 
was challenged when \citet{1999Sci...283.1290C}
introduced a non-equilibrium kinetic theory by which carbon can 
condense thermally in gas having more abundant oxygen, opening the 
possibility of C condensation throughout the O-rich core. That 
initial theory has been amplified by several studies 
\citep{2001ApJ...562..480C,2003ApJ...594..312D,2006ApJ...638..234D,
2011NewAR..55..155C, 2013ApJ...762....5C}.
Within their chemical model \citet{2009ApJ...703..642C,2010ApJ...713....1C}
raised questions about the most effective way of producing $\ctwo$\ molecules, 
which is the first step to linear carbon chains. They discovered that 
the neutral-neutral reaction \coctoctwo\ has a reasonably large cross section 
$\langle \sigma v \rangle = k_{CO,C} = 8.6 \times 10^{-14}$ cm$^3$ s$^{-1}$
at temperatures near 5000K despite being endothermic by 4.8eV. 
Although the CO target abundance for production of $\ctwo$\ in this 
manner must be calculated, they stated that in their model 
at $T = 5000$ K it created $\ctwo$\ much more rapidly than the slow radiative 
reaction \twoctoctwo. Sensing high importance of this reaction for astronomy, 
they claimed in their list of conclusions; 2. A new pathway to the 
formation of carbon chains is active in the O-rich mass zone of the 
unmixed ejecta and is identified as the CO conversion to $\ctwo$\ via 
collisions with C. \citet{2009ApJ...703..642C} were unable to evaluate 
the quantitative consequences because their reaction network 
terminated at $\cthree$\ and did not include the formation of linear 
carbon chains $\cenn$\ and their isomerization to carbon rings. 
\citet{2010ApJ...713....1C} subsequently did include 
linear chains to $n=10$, which they took to isomerize instantly 
to ringed structure, following the published 
model \citet{1999Sci...283.1290C, 2001ApJ...562..480C}.

Testing their calculated abundances for $\cenn$\ is but a secondary 
goal of the present study. Its overriding goal is to point out 
abundance differences of several orders of magnitude for the CO 
and $\ctwo$\ molecules between the computations of these two groups 
and to resolve those differences if possible. This test has 
important consequences for condensation of carbon dust, changing 
prior expectations if the production of $\ctwo$\ owing to the reaction 
\twoctoctwo\ severely underestimates the subsequent abundances of 
linear carbon chains. \citet{2010ApJ...713....1C} harbored that 
expectation owing to their belief that much more $\ctwo$\ is 
made earlier near $T = 5000$ K and survives. Dust created by 
supernova expansions is currently under intense study owing to three 
types of astronomical observations: (1) of dust observed in single 
supernova remnants; (2) of dust observed in early low-metallicity 
galaxies; (3) of supernova-condensed carbon dust (SUNOCONs) 
extracted from meteorites. Each of these topics depends sensitively on 
how much carbon, both numbers and sizes, is able to condense in 
cooling SNII interiors. Therefore we study this competition carefully.

Crucial to this task is the lifetime $\tau_{CO}$ of CO molecules. 
Thermal dissociation of CO is dominated by thermal photons because 
radiative association of C and O dominates other reactions for 
the formation of CO at the densities within SNII. Each reaction 
and its inverse is subject to the quantum principle of detailed 
balance. The thermal photodissociation lifetime $\tau_\gamma$ is 
calculated from detailed balance with the radiative association reaction, as in 
Section 2.1 of \citet{2001ApJ...562..480C}. Because of the large 11.1eV 
binding energy of the CO molecule, $\tau_\gamma$ is very temperature sensitive. 
For readers’ numerical ease we will tabulate $\tau_\gamma$ at selected key 
temperatures in Table 1 to follow. The flux and spectrum of newly 
injected Compton electrons \citep{1991ApJ...375..221C} yields the 
lifetime $\tau_e$ of CO molecules against inelastic scattering dissociation by 
Compton electrons caused by the $^{56}$Co radioactivity. In a gas of pure 
CO the mean energy per ion pair is defined as the energy of primary 
electrons divided by the number of pairs produced. 
\citet{1994ApJ...435..909L} calculated the mean energy per ion pair 
in pure CO gas and obtained the result $\Delta E = 32.3$ eV deposited 
per dissociated CO pair, agreeing with the measurement of 32.2 eV 
by \citet{1968...Klots}. So the efficiency of energetic electrons for 
dissociating CO seems well established. A lifetime $\tau_{CO}$ near one 
week is typical in SN 1987A, but would be longer in SNII 
synthesizing less $^{56}$Co and at times greater than about 8 months.

@@

Because our goal is to study the chemistry, we can take a very simple physical
model of the expansion; namely, temperature
$T = 3800 K / (t / 100 {\rm d}) =
3.30 \times 10^{10} / t({\rm s})\ {\rm K}$,
where t is the time elapsed since core collapse.
We chooose $n_O = 10^{10}$ cm$^{-3}$
and $n_C = 10^9$ cm$^{-3}$
at the starting time $t = t(6000 {\rm K})$ 
at $T=6000$ K for our chemical network . From our model choice for $T(t)$ we get
$t(6000K) = 5.47 \times 10^6$s. At
subsequent times $n_O(t) = 10^{10}{\rm cm}^{-3}(t/t(6000 {\rm K}))^{-3}$
owing to homologous expansion.
Let $N$ be the number of any specific molecular species
in a comoving, expanding, initially 1 cm$^3$ volume at 6000 K.
The only change of $N$ during
expansion occurs through chemical reactions. $N$ may be expressed as atom
fraction $Y$ of the initial total number
$N_O + N_C = 1.10 \times 10^{10}$ atoms. We
intentionally choose an O-rich interior having $N_O/N_C = 10$,
so that no carbon at
all would be able to condense if that interior were governed by chemical
equilibrium.

We use an intentionally limited set of chemical species because our goal is
to study the controversy over the correct 
carbon chemistry pathway to the ringed isomers.
We limit the present study to C, O, CO, $\ctwo$, $\cthree$, ... $\ceight$\ and
$\ceightr$, the ringed isomer of $\ceight$, to which we give a lifetime
$\tau_8^r = 10$ s against thermal isomeric transition
from linear $\ceight$ to ringed $\ceightr$.
In a sense the $n = 8$ ring is a dummy standing for all rings; but it is also
reasonable in being the smallest ring that is widely expected 
(e.g. \citealt{doi:10.1021/j100374a025}).
Those ringed molecules are the seeds for carbon
growth because their oxidation rates are much smaller than oxidation rates of
linear $\cenn$
whereas their C capture rates are fast. 
For purposes of this study we take ringed $\ceightr$ to be
indestructible, simply integrating its rate of production.
%We cannot allege that $\ceight$\ is the key isomerizing chain, nor that
%$\tau_8^r = 10$ s is its lifetime, but those assumptions are adequate for
%comparing models for formation for $\ctwo$.
Our strategy is
to compute the number $\ceightr$ remaining after expansion from two differing
sources of $\ctwo$\ initiating the linear $\cenn$
chains. Those two source reactions are
\begin{itemize}
\item \twoctoctwo\ \citep{1999Sci...283.1290C}.
\item \coctoctwo\ \citep{2009ApJ...703..642C,2010ApJ...713....1C}.
\end{itemize}
We take our reaction rates from the rate tables in
\citet{2009ApJ...703..642C,2010ApJ...713....1C}.
The rate for the second reaction is indeed
much greater than that of the first reaction above $T=3000$ K;
but it becomes increasingly the smaller of the two below $T=2500$ K.
We include the thermal inverse reaction of every reaction,
which we calculate as in Eq.(3) of \citet{2001ApJ...562..480C}.
See Table 1 below for sample thermal dissociation rates for CO molecules.
We also include in our network dissociation of CO by Compton electrons.
CO is the only molecule for which electron dissociation can be the dominant
destruction mechanism. For that rate we use
$\tau_e (s) = 10^5 \exp( (t-10^6{\rm s}) / 111 {\rm d} )$,
not fit to any specific model but with plausibility for SN1987A.
It lengthens to $\tau_e = 10^6$ s near eight months.
Because of their helpful cataloging of reaction rate tables,
and in order to compare results without rate differences, we take the
rates as given in the tables of
\citet{2009ApJ...703..642C,2010ApJ...713....1C}.
As one example, we take the rate for \twoctoctwo\
as the rate given by RA4 in Table 5 in \citet{2009ApJ...703..642C}, p. 649.

@@

Evidently the abundance of CO within that zone during expansion attempts to
balance these creation and destruction effects, doing so exactly at the
time $t = t_{max}$ of maximum $n_{CO}$. \citet{2013ApJ...762....5C}
has discussed the shape of the function
$n_{CO}$(t). His Eq.(1), which is valid at constant density,
approximates the growth of $n_{CO}$ by its leading terms:
\begin{equation}
\frac{dn_{CO}}{dt} = n_C n_O k_{CO} - \frac{n_{CO}}{\tau_{CO}} = 0
\label{eq:dncodt}
\end{equation}
at $t = t_{max}$.
The maximum abundance reached by CO is in cgs units
\begin{equation}
n_{CO}(t_{max}) = n_C n_O k_{CO} \tau_{CO} 
\label{eq:comax}
\end{equation}

@@

Figure \ref{fig:yi}
shows the abundances $Y_i$ of each species in our small network as a
function of time $t-t(6000 {\rm K})$ after $T=6000$ K.
From Table \ref{tab:quantities},
start time is $t(6000K) =5.47 \times 10^6$ s.
To display each $Y_i$ on a figure with
reasonable ordinate resolution, we have scaled each $Y_n$\ by the factor
stated in the box. Many features are noteworthy: (1) $Y_C$ and $Y_O$ 
are constant
because their small depletion is negligible on the scale shown;
(2) maxima of $Y_{CO}$ and $Y_2$\ occur at almost exactly the same time
$t-t(6000K)= 4 \times 10^6$ s, as was also seen in Fig. \ref{fig:ncoc2};
(3) $\cthree$\ has very small abundance, about $10^{-7}$ of $Y_2$\
at the $Y_2$\ maximum, although much later $Y_3$\ slowly grows modestly
relative to much more abundant $\ctwo$; (4) the rise shapes of $Y_4$
through $Y_8$\ are very similar because they are linked by a near steady
state; (5) the ringed carbon $Y(\ceightr$) has similar abundance shape vs. time,
but notice that it is actually much more abundant than lineaer $C_{7-8}$ and
because it accumulates from isomeric transitions of $\ceight$\
and unlike $\cenn$\ does not
suffer from fast oxidation \citep{1999Sci...283.1290C}. Each of these
features is understandable in terms of the flows into and out of each species.

One sees from Fig. \ref{fig:yi} that the \coctoctwo\ reaction plays no
role in $\ceightr$\ production from the fact that $\ceightr$\ rises only for
$t - t(6000 {\rm K}) > 1.5 \times 10^7$ s, corresponding to $T < 2000$ K,
despite $\ctwo$\ and $\cthree$\ rising at much earlier times.
Whatsoever $\ctwo$\ is made earlier plays no role in $\ceightr$\ production
because the ejection of C atoms from $\cenn$\ at higher $T$ prevents the
flow to $\ceight$.  Only at $T < 2000$ K do those photodissociation reactions
become so slow that $\ceight$\ can grow, which it does from the $\ctwo$\
recently
formed by the \twoctoctwo\ reaction near 2000 K.  Fig. \ref{fig:yi} also shows
that $Y(\ceightr)$
grows quickly to $10^{-16}$.  Since available C is $Y_C = 0.1$,
the ratio gives $10^{-15}$ C/ring.  This is enough to grow very large
graphite.

Figure \ref{fig:cn} shows the shape of $\cenn$\ vs. $n$ for $\cthree$\ to
$\ceight$\ at two different temperatures near 2000 K. As temperature declines,
the curve flattens because photoejection from $\cenn$\ by thermal photons
(or vibrations) weakens. These abundance ratios are almost in a steady state,
but that steady state changes slightly as $T$ falls. These patterns show
almost equal values for $Y(\cthree)$ because that curve is relatively flat
with time near $t-t(6000K) = 10^7$ s (Fig. \ref{fig:yi}).
Table \ref{tab:flows} lists our computed reaction flows both in and out of
$\csix$\ at $T = 2031$ K and $T = 1802$ K.
The reactions are specified there in the
compact notation target(in,out)residual. 
We tabulate the flow magnitudes for $\csix$ to aid readers checking our
numerical results. Fig. \ref{fig:cn} and
Table \ref{tab:flows} are made to be studied together.
Note that the abundances of $\ceight$\ are very small:
$Y(\ceight)= 10^{-36}$ at $T = 1802$ K.
The isomeric transition from linear to
ringed $\ceight$, for which we estimate a lifetime $\tau = 10$ s,
provides the nucleations for graphite growth.

We call attention to several conclusions to be drawn from Table \ref{tab:flows}.
Firstly, the strongest flows by a wide margin involving $\csix$\ are
\cfivectocsix,
which is very fast ($k = 3 \times 10^{-10} {\rm cm}^3 {\rm s}^{-1}$)
because vibrational excitation of $\csix$
obviates the need for creating a photon \citep{1999Sci...283.1290C},
and its inverse reaction which ejects a C atom. Furthermore, those two flows
are equal to each other to three significant figures, illustrating the near 
steady
state of the abundance pattern. Secondly, thermal dissociation of $\csix$\ is
very much faster than its oxidation, showing the small effect of oxidation
on the abundance pattern. The destroying flows from $\csix$\ to $\cfive$
stand in the approximate ratio
${\rm C}_6(\gamma,{\rm C})/{\rm C_6}({\rm O},{\rm CO}) = 5 \times 10^5$
at $T = 2031$ K and $2 \times 10^4$ at $T = 1802$ K.
Dissociation by thermally excited vibrations dominates oxidation. What we here
label ($\gamma,{\rm C}$) is actually radiationless. Note that oxidation of
$\cenn$ molecules is not faster than C-ejection reactions. This unusual
situation occurs because the C-capture reactions proceed by exciting 
vibrations of the $\csix$ molecule. That residual vibrational energy makes
capture reactions fast rather than slow, but also makes the inverse C-ejection
reactions caused by thermally excited vibrations to be much faster. Detailed
balance gets the ratio right. The smaller value for that ratio at $T=1802$ K
occurs because oxidation maintains its effectiveness as $T$ drops but
thermal disruption does not, being much more temperature sensitive.
For this reason the abundance pattern is flatter and $\csix$\ is approximately
3000 times more abundant at $T=1802$ K than at $T=2031$ K.
It is for that reason that destruction flows in Table \ref{tab:flows}
for $\csix$\ are
larger at the smaller temperature.
The steady state shifts to new ratios as $T$ falls.
Thirdly, calculation of the rates will be illustrated for sake of clarity
by the flow
\csixotococfive.
From Fig. \ref{fig:cn}
one sees that at $T = 2031$ K the abundance $Y(\csix) = 10^{-33.3}$.
The value of $Y(O) = 10^{10} cm^{-3} / 1.1 \times 10^{10} cm^{-3} = 0.909$,
so the flow per atom per
second (e.g., Eq. \ref{eq:dycodt}) is
$dY/dt= Y(\csix)Y(O)n(2031K)k(\csix + {\rm O})$.
Expansion from $T=6000$ K to $T=2031$ K has diluted the total number
density $n$ to $1.1 \times 10^{10}{\rm cm}^{-3}
(2031 {\rm K}/6000 {\rm K})^3 = 4.27 \times 10^8{\rm cm}^{-3}$.
The fast oxidation reaction rate factor is
$k(\csix + {\rm O}) = 3 \times 10^{-10} {\rm cm}^3{\rm s}^{-1}$.
Gathering factors yields approximately $4 \times 10^{-35}$ per atom
per second
in good approximation to the Table \ref{tab:flows}
flow entry $3.61 \times 10^{-35}$ per atom per second.

@@

\begin{table}
\begin{center}
\caption{Values of $\tau_\gamma$(CO) and related quantities at
selected time $t$ and temperature $T$}
\label{tab:quantities}
\begin{tabular}{ccccccc}
\tableline\tableline
$t$ (s) & $\left[t / t(6000 {\rm K}) \right]^{-3}$ & $T$ (K) & $\tau_\gamma$ (s) & $k_{CO}$ (cm$^{-3}$ s$^{-1}$) & $\ncogeq$ (cm$^{-3}$) \\
\tableline
$5.47 \times 10^6$ &
  1 &
  6000 &
  $4.22 \times 10^{-2}$ &
  $3.13 \times 10^{-17}$ &
  $1.32 \times 10^1$ \\
$5.97 \times 10^6$ &
  0.769 &
  5500 &
  $3.54 \times 10^{-1}$ &
  $2.99 \times 10^{-17}$ &
  $6.27 \times 10^1$ \\
$6.57 \times 10^6$ &
  0.578 &
  5000 &
  $4.46 \times 10^{0}$ &
  $2.83 \times 10^{-17}$ &
  $4.24 \times 10^2$ \\
$7.30 \times 10^6$ &
  0.422 &
  4500 &
  $9.70 \times 10^{1}$ &
  $2.67 \times 10^{-17}$ &
  $4.61 \times 10^3$ \\
$8.21 \times 10^6$ &
  0.296 &
  4000 &
  $4.43 \times 10^{3}$ &
  $2.48 \times 10^{-17}$ &
  $9.67 \times 10^4$ \\
$9.38 \times 10^6$ &
  0.198 &
  3500 &
  $5.86 \times 10^{5}$ &
  $2.28 \times 10^{-17}$ &
  $5.26 \times 10^6$ \\
$1.09 \times 10^7$ &
  0.125 &
  3000 &
  $3.77 \times 10^{8}$ &
  $2.05 \times 10^{-17}$ &
  $1.21 \times 10^9$ \\
\tableline
\end{tabular}
\end{center}
\end{table}

@@

\bibitem[{Takai {et~al.}(1990)Takai, Lee, Halicioglu, \&
  Tiller}]{doi:10.1021/j100374a025}
Takai, T., Lee, C., Halicioglu, T., \& Tiller, W.~A. 1990, The Journal of
  Physical Chemistry, 94, 4480
